
---
title: "LendingClubAnalysis"
author: "Steve Isaacs"
date: "1/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(formattable)
library(ggthemr)
```


Business Problem: Assessing credit worthiness of customers in Lending Club and their probability of defaulting.

Data set: https://www.kaggle.com/wendykan/lending-club-loan-data

### 1.Introduction to the Business Problem

Lending Club is the world’s largest online marketplace connecting borrowers and investors.They operate at a lower cost than traditional bank lending programs and pass the savings on to borrowers in the form of lower rates and to investors in the form of solid returns.
Source: https://www.lendingclub.com/ 

THis project aims to analyse Lending Club's issued loans over 9 year period (2007-2015) and zdentify early indicators that could predict a customer's probability of defaulting. The insight obtained could help to drive differentiated credit approval process for different customer segment (e.g faster approval time, minimal security for low risk customers)

<br>
### 2.Process to solve the Business Problem

We would suggest the following process
1. Understand Lending Club Business Model & Data Set
2. Clean the data, identify relevant attributes/columns & separate training and testing dataset
3. Generate Hypothesis 
4. Apply Dimensionality Reduction on Hypothesis
5. Apply Segmentation & Classification Tree on Training Dataset
6. Apply Algorithm on Testing Dataset (Iterate if required)
7. Capture insights & Summarise

<br>

### 3.Key Summary of Data

```{r eval = FALSE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='markup'}
Projectdata<- read.csv(file = "Data/loan.csv", header = TRUE, sep=",")
print(colnames(Projectdata))

#Reduce Projectdata set into smaller dataframe name Loandata of identified Relevant Variables

Loandata <- Projectdata[, c("addr_state", "annual_inc", "annual_inc_joint","application_type", "delinq_2yrs", "emp_length" ,"funded_amnt" ,"funded_amnt_inv", "grade", "home_ownership", "id" , "loan_amnt" , "loan_status", "member_id" , "mths_since_last_delinq" , "open_acc" , "out_prncp" ,"out_prncp_inv", "pub_rec", "purpose" , "revol_bal" , "revol_util" , "sub_grade", "term", "total_acc" ,"verification_status_joint", "verification_status")]

install.packages("DescTools")
library(DescTools)
options(scipen=1000)

#look at distribution of LoanAmount
Desc(Loandata$loan_amnt, main="Loan Amount",plotit = TRUE)
Desc(Loandata$loan_status, main= "Loan Status", plotit = T)

#Visualise Loan Amount by Status
#box_status <- ggplot(Loandata, aes(loan_status, loan_amnt))
#box_status + geom_boxplot(aes(fill = loan_status)) + theme(axis.text.x = element_blank()) + labs(list(title = "Loan amount by status", x = "Status",y = "Amount"))

```
<br>

### 4.Dimensionality Reduction
```{r eval = FALSE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='markup'}
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.5)
options(knitr.kable.NA = '')

dformat <-function(df) {
  if (class(df) != "data.frame")
    df <- as.data.frame(df)
  x <- lapply(colnames(df), function(col) {
    if (is.numeric(df[, col]))
      color_bar(rgb(238, 238, 238, max=255), min=0.1, na.rm=TRUE)
    else
      formatter("span")
  })
  names(x) <- colnames(df)
  formattable(df, x)
}

# SET UP OF ALL THE INPUTS FOR THIS PART

# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
MIN_VALUE=0.1

# Please enter the maximum number of observations to show in the report and slides 
# (DEFAULT is 15. If the number is large the report and slides may not be generated - very slow or will crash!!)
max_data_report = 10

# Please ENTER then original raw attributes to use. 
# Please use numbers, not column names! e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
factor_attributes_used= c(2:30)

# Please ENTER the selection criterions for the factors to use. 
# Choices: "eigenvalue", "variance", "manual"
factor_selectionciterion = "eigenvalue"

# Please ENTER the desired minumum variance explained 
# (ONLY USED in case "variance" is the factor selection criterion used). 
minimum_variance_explained = 65  # between 1 and 100

# Please ENTER the number of factors to use 
# (ONLY USED in case "manual" is the factor selection criterion used).
manual_numb_factors_used = 15

# Please ENTER the rotation eventually used (e.g. "none", "varimax", "quatimax", "promax", "oblimin", "simplimax", and "cluster" - see help(principal)). Defauls is "varimax"
rotation_used="varimax"

factor_attributes_used = unique(sapply(factor_attributes_used,function(i) min(ncol(Loandata), max(i,1))))
LoandataFactor=Loandata[,factor_attributes_used]
LoandataFactor <- Loandata <- data.matrix(LoandataFactor)

#Check Correlations
thecor = round(cor(LoandataFactor),2)

minval <- 0.1
thecor_thres <- thecor
thecor_thres[abs(thecor_thres) < minval]<-"-"

colnames(thecor_thres)<-colnames(LoandataFactor)
rownames(thecor_thres)<-colnames(LoandataFactor)

dformat(thecor_thres)

```

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='markup'}

library(psych)
library(FactoMineR)
library(magrittr)

# Here is how the `principal` function is used
UnRotated_Results<-principal(LoandataFactor, nfactors=ncol(LoandataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Component",1:ncol(UnRotated_Factors),sep=" ")

# Here is how we use the `PCA` function 
Variance_Explained_Table_results<-PCA(LoandataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table


rownames(Variance_Explained_Table) <- paste("Component", 1:nrow(Variance_Explained_Table))
colnames(Variance_Explained_Table) <- c("Eigenvalue", "Pct of explained variance", "Cumulative pct of explained variance")

#Look at the Variance Explained Table
dformat(round(Variance_Explained_Table, 2))

#Look at EigenValues
eigenvalues  <- Variance_Explained_Table[, "Eigenvalue"]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
#ggplot(melt(df, id="components"), aes(x=components, y=value, colour=variable)) + geom_line()
#c3(melt(df, id="components"), x="components", y="value", group="variable") %>% c3_line('spline')

```

<br>
### 5.Clustering and Segmentation

<br>
### 6. Results 


